---
title: "The SABRhood Daily"
---

```{r}
#| echo: false
#| message: false
#| warning: false

library(tidyverse)
library(glue)
library(knitr)

newsletter_path <- file.path("data", "newsletter_data", "newsletter_today.rds")

if (!file.exists(newsletter_path)) {
  stop("Missing newsletter_today.rds. Run source('scripts/newsletter/update_newsletter.R') first.")
}

daily <- readRDS(newsletter_path)
```

```{r}
#| echo: false
#| message: false
#| warning: false

escape_html <- function(x) {
  x <- as.character(x)
  x <- gsub("&", "&amp;", x, fixed = TRUE)
  x <- gsub("<", "&lt;", x, fixed = TRUE)
  x <- gsub(">", "&gt;", x, fixed = TRUE)
  x <- gsub('"', "&quot;", x, fixed = TRUE)
  x
}

null_or_na <- function(x) {
  is.null(x) || length(x) == 0 || all(is.na(x))
}

safe_text <- function(x, fallback = "—") {
  if (null_or_na(x)) return(fallback)
  escape_html(x[[1]])
}

fmt_board_field <- function(value, col) {
  if (null_or_na(value)) return("—")

  if (col == "IP") {
    return(sprintf("%.1f", suppressWarnings(as.numeric(value))))
  }

  if (col %in% c("hot_score", "best_z", "z_score", "pitcher_game_score", "hitter_game_score", "diff")) {
    return(as.character(round(suppressWarnings(as.numeric(value)), 1)))
  }

  if (is.numeric(value)) {
    return(as.character(round(value, 1)))
  }

  escape_html(value)
}

pretty_label <- function(x) {
  dplyr::recode(
    x,
    "team" = "Team",
    "H" = "H",
    "HR" = "HR",
    "TB" = "TB",
    "RBI" = "RBI",
    "SO" = "K",
    "whiffs" = "Whiffs",
    "IP" = "IP",
    "best_category" = "Type",
    "best_stat" = "Stat",
    "direction" = "Move",
    "stat" = "Stat",
    "pitch_type" = "Pitch",
    .default = stringr::str_to_title(gsub("_", " ", x))
  )
}

render_img <- function(path, alt = "") {
  if (!null_or_na(path)) {
    path <- as.character(path[[1]])

    img_id <- paste0(
      "img-",
      stringr::str_replace_all(
        stringr::str_to_lower(basename(path)),
        "[^a-z0-9]+",
        "-"
      )
    )

    cat(
      '<a class="daily-img-link" href="#', img_id, '">',
      '<img src="', escape_html(path), '" alt="', escape_html(alt), '">',
      '</a>',
      '<div id="', img_id, '" class="daily-lightbox">',
      '<a class="daily-lightbox-close" href="#_">×</a>',
      '<img src="', escape_html(path), '" alt="', escape_html(paste(alt, "expanded")), '">',
      '</div>',
      sep = ""
    )
  }
}

render_evidence_strip <- function(items) {
  if (is.null(items) || length(items) == 0) {
    return(invisible(NULL))
  }

  cat('<div class="daily-evidence-strip">', sep = "")

  for (item in items) {
    cat(
      '<div class="daily-evidence-pill">',
      '<span>', safe_text(item$label), '</span>',
      '<strong>', safe_text(item$value), '</strong>',
      '</div>',
      sep = ""
    )
  }

  cat('</div>', sep = "")
}

render_mini_board <- function(tbl, name_col, sub_cols = NULL, value_col, value_suffix = "", n = 6) {
  if (is.null(tbl) || nrow(tbl) == 0 || !name_col %in% names(tbl) || !value_col %in% names(tbl)) {
    cat('<p class="daily-small-note">No qualifying rows available.</p>', sep = "")
    return(invisible(NULL))
  }

  tbl <- tbl %>% slice_head(n = n)

  cat('<div class="daily-mini-board">', sep = "")

  for (i in seq_len(nrow(tbl))) {
    main <- tbl[[name_col]][i]
    value <- tbl[[value_col]][i]

    subline <- ""
    if (!is.null(sub_cols)) {
      subline <- paste(
        purrr::map_chr(sub_cols, function(col) {
          if (!col %in% names(tbl)) return("")
          paste0(pretty_label(col), ": ", fmt_board_field(tbl[[col]][i], col))
        }) %>% discard(~ .x == ""),
        collapse = " • "
      )
    }

    cat(
      '<div class="daily-board-row">',
      '<span class="daily-rank">', i, '</span>',
      '<div class="daily-board-main">',
      '<strong>', escape_html(main), '</strong>',
      '<span>', subline, '</span>',
      '</div>',
      '<div class="daily-board-value">', fmt_board_field(value, value_col), value_suffix, '</div>',
      '</div>',
      sep = ""
    )
  }

  cat('</div>', sep = "")
}
```

<div class="daily-paper">

```{r}
#| echo: false
#| results: asis

cat(
  '<section class="daily-masthead">',
  '<div class="daily-meta">',
  '<span class="daily-chip">', safe_text(daily$weekday), ' Edition</span>',
  '<span class="daily-chip">', safe_text(daily$edition_type), '</span>',
  '</div>',
  '<h1>The SABRhood Daily</h1>',
  '<p>', safe_text(daily$lead_blurb), '</p>',
  '</section>',
  sep = ""
)
```

<div class="daily-grid">

<div class="daily-stack">

<section class="daily-section daily-lead-card">
  <span class="daily-kicker">Lead Board</span>
  <h2>`r daily$headline`</h2>
  <p>`r daily$deck`</p>
</section>

<section class="daily-section">
  <span class="daily-kicker">Featured Players</span>
  <h2>Hot bat and hot arm</h2>

  <div class="daily-card-grid">

```{r}
#| echo: false
#| results: asis

h <- daily$featured_hitter_card

if (!is.null(h)) {
  cat('<article class="daily-player-card">', sep = "")
  cat('<span class="daily-kicker">', safe_text(h$tag), '</span>', sep = "")
  cat('<h3>', safe_text(h$headline), '</h3>', sep = "")
  cat('<p class="daily-subline">', safe_text(h$team), ' • ', safe_text(h$trend_stat), ' movement</p>', sep = "")
  cat('<p>', safe_text(h$summary), '</p>', sep = "")

  render_evidence_strip(h$evidence)

  cat('<p class="daily-small-note">', safe_text(h$season_line), '</p>', sep = "")

  render_img(h$chart_path, paste("Rolling trend chart for", h$player))

  cat('</article>', sep = "")
}

p <- daily$featured_pitcher_card

if (!is.null(p)) {
  cat('<article class="daily-player-card">', sep = "")
  cat('<span class="daily-kicker">', safe_text(p$tag), '</span>', sep = "")
  cat('<h3>', safe_text(p$headline), '</h3>', sep = "")
  cat('<p class="daily-subline">', safe_text(p$team), ' • pitcher trend card</p>', sep = "")
  cat('<p>', safe_text(p$summary), '</p>', sep = "")

  render_evidence_strip(p$evidence)

  cat('<p>', safe_text(p$result_line), '</p>', sep = "")
  cat('<p>', safe_text(p$whiff_line), '</p>', sep = "")
  cat('<p><strong>Pitch mix:</strong> ', safe_text(p$pitchmix_note), '</p>', sep = "")

  render_img(p$chart_path, paste("Pitch mix chart for", p$pitcher))

  cat('</article>', sep = "")
}
```

  </div>
</section>

<section class="daily-section">
  <span class="daily-kicker">Pitch Lab</span>
  <h2>Pitch of the Day</h2>

```{r}
#| echo: false
#| results: asis

pod <- daily$pitch_of_day

if (!is.null(pod)) {
  cat('<div class="daily-player-card">', sep = "")
  cat('<span class="daily-kicker">', safe_text(pod$tag), '</span>', sep = "")
  cat('<h3>', safe_text(pod$title), '</h3>', sep = "")
  cat('<p class="daily-subline">', safe_text(pod$pitcher), ' • ', safe_text(pod$team), '</p>', sep = "")
  cat('<p>', safe_text(pod$note), '</p>', sep = "")
  render_evidence_strip(pod$evidence)
  cat('</div>', sep = "")
} else {
  cat('<p class="daily-small-note">No pitch qualified for today’s Pitch of the Day.</p>', sep = "")
}
```

</section>

<section class="daily-section">
  <span class="daily-kicker">Split Lens</span>
  <h2>Where the profile bends</h2>

```{r}
#| echo: false
#| results: asis

split_items <- list(
  daily$split_spotlight$hitter,
  daily$split_spotlight$pitcher
)

split_items <- split_items[!purrr::map_lgl(split_items, is.null)]

if (length(split_items) > 0) {
  cat('<div class="daily-card-grid">', sep = "")

  for (item in split_items) {
    cat('<article class="daily-player-card">', sep = "")
    cat('<span class="daily-kicker">', safe_text(item$type), '</span>', sep = "")
    cat('<h3>', safe_text(item$player), '</h3>', sep = "")
    cat('<p class="daily-subline">', safe_text(item$comparison), '</p>', sep = "")
    cat('<p>', safe_text(item$note), '</p>', sep = "")

    render_evidence_strip(list(
      list(label = "Stat", value = item$stat),
      list(label = "Diff", value = round(item$diff, 3)),
      list(label = "View", value = "Split")
    ))

    cat('</article>', sep = "")
  }

  cat('</div>', sep = "")
} else {
  cat('<p class="daily-small-note">No split spotlight qualified today.</p>', sep = "")
}
```

</section>

<section class="daily-section daily-chart-card">
  <span class="daily-kicker">Visual Read</span>
  <h2>`r daily$rolling_ops_comparison$title`</h2>
  <p>`r daily$rolling_ops_comparison$note`</p>

```{r}
#| echo: false
#| results: asis

if (!is.null(daily$rolling_ops_comparison)) {
  render_img(daily$rolling_ops_comparison$path, "Rolling OPS comparison")
}
```

</section>

<section class="daily-section daily-chart-card">
  <span class="daily-kicker">Chart of the Day</span>
  <h2>`r daily$chart_of_day$title`</h2>
  <p>`r daily$chart_of_day$note`</p>

```{r}
#| echo: false
#| results: asis

if (!is.null(daily$chart_of_day)) {
  render_img(daily$chart_of_day$path, "Chart of the day")
}
```

</section>

<section class="daily-section">
  <span class="daily-kicker">Signal Boards</span>
  <h2>Who else is moving?</h2>

  <div class="daily-card-grid">

  <article class="daily-player-card">
    <span class="daily-kicker">Pitcher Trends</span>
    <h3>Hot arm board</h3>

```{r}
#| echo: false
#| results: asis

render_mini_board(
  tbl = daily$pitcher_hot_board,
  name_col = "pitcher",
  sub_cols = c("team", "best_category", "best_stat"),
  value_col = "hot_score",
  n = 7
)
```

  </article>

  <article class="daily-player-card">
    <span class="daily-kicker">Hitter Trends</span>
    <h3>Movement board</h3>

```{r}
#| echo: false
#| results: asis

if (!is.null(daily$trenders) && nrow(daily$trenders) > 0) {
  movement_board <- daily$trenders %>%
    arrange(desc(abs(z_change))) %>%
    distinct(player, .keep_all = TRUE) %>%
    mutate(
      z_score = round(z_change, 2)
    )

  render_mini_board(
    tbl = movement_board,
    name_col = "player",
    sub_cols = c("team", "stat", "direction"),
    value_col = "z_score",
    n = 8
  )
} else {
  cat('<p class="daily-small-note">No hitter movement rows available.</p>', sep = "")
}
```

  </article>

  </div>
</section>

<section class="daily-section">
  <span class="daily-kicker">Trend Queue</span>
  <h2>Names to watch</h2>

```{r}
#| echo: false
#| results: asis

if (!is.null(daily$watch_list) && nrow(daily$watch_list) > 0) {
  watch_tbl <- daily$watch_list %>% slice_head(n = 6)

  cat('<div class="daily-card-grid">', sep = "")

  for (i in seq_len(nrow(watch_tbl))) {
    cat('<article class="daily-player-card">', sep = "")
    cat('<span class="daily-kicker">', escape_html(watch_tbl$type[i]), '</span>', sep = "")
    cat('<h3>', escape_html(watch_tbl$name[i]), '</h3>', sep = "")
    cat('<p class="daily-subline">', escape_html(watch_tbl$team[i]), ' • ', escape_html(watch_tbl$tag[i]), '</p>', sep = "")
    cat('<p>', escape_html(watch_tbl$reason[i]), '</p>', sep = "")
    cat('</article>', sep = "")
  }

  cat('</div>', sep = "")
} else {
  cat('<p class="daily-small-note">No watch-list rows available.</p>', sep = "")
}
```

</section>

<section class="daily-section">
  <span class="daily-kicker">Method</span>
  <h2>Trend key</h2>

```{r}
#| echo: false
#| results: asis

if (!is.null(daily$trend_legend) && nrow(daily$trend_legend) > 0) {
  cat('<div class="daily-card-grid">', sep = "")

  for (i in seq_len(nrow(daily$trend_legend))) {
    cat('<article class="daily-player-card">', sep = "")
    cat('<span class="daily-kicker">Signal</span>', sep = "")
    cat('<h3>', escape_html(daily$trend_legend$tag[i]), '</h3>', sep = "")
    cat('<p>', escape_html(daily$trend_legend$meaning[i]), '</p>', sep = "")
    cat('</article>', sep = "")
  }

  cat('</div>', sep = "")
} else {
  cat('<p class="daily-small-note">Trend key unavailable.</p>', sep = "")
}
```

</section>

</div>

<aside class="daily-stack">

<section class="daily-section">
  <span class="daily-kicker">Last Night</span>
  <h2>Top bats</h2>

```{r}
#| echo: false
#| results: asis

render_mini_board(
  tbl = daily$last_night$hitters,
  name_col = "player",
  sub_cols = c("team", "H", "HR", "TB", "RBI"),
  value_col = "hitter_game_score",
  n = 6
)
```

</section>

<section class="daily-section">
  <span class="daily-kicker">Last Night</span>
  <h2>Top arms</h2>

```{r}
#| echo: false
#| results: asis

render_mini_board(
  tbl = daily$last_night$pitchers,
  name_col = "pitcher",
  sub_cols = c("team", "IP", "SO", "whiffs"),
  value_col = "pitcher_game_score",
  n = 6
)
```

</section>

</aside>

</div>

</div>